### 2022-06-13 paper.tex
minor stuff + discussion draft

\begin{figure}
    A\\
    \includegraphics[width=.7\columnwidth]{part_leonie_git/leonie_plots/corr_methylation_fraction_4kbwindow.pdf} \\
    B\\
    \includegraphics[width=.7\columnwidth]{part_leonie_git/leonie_plots/corr_shrunken_residual_4kbwindow.pdf} \\
    C\\
    \includegraphics[width=.95\columnwidth]{part_leonie_git/leonie_plots/comparison_meth_RNA.pdf}
    \caption{\small \textbf{Correlation of DNA methylation and gene expression}.
    \textbf{(A)} Correlation of each gene's RNA expression values to the the methylation fractions that were measured at promoters (y-axis), or to methylation fractions of the nearest VMR (x-axis). Promoters are defined as intervals \textpm2~kb around the TSS.
    \textbf{(B)} Same as in (A), but using mean shrunken residuals as a measure of methylation.
    \textbf{(C)} Mean smoothed methylation near the gene Ube2l6 (ENSMUSG00000027078, marked in yellow in A \& B). Cells are assigned to 4 groups based on Ube2l6 expression (group 0: cells with no Ube2l6 expression, groups 1-3: cells that express Ube2l6, divided into three equally large groups with group 3 having the highest expression). The mean methylation level of each group was smoothed with a tricube kernel of bandwidth 2000.
    Data from \citet{argelaguet2019gastru}.}
    \label{figure:correlation}
\end{figure}

@@

To assess the impact of our methods on dimensionality reduction and the ability to distinguish different (sub)types of cells, we used the data set from \citet{luo2017single}. This data set consists of over three thousand \todo{Give exact number.} single-cell CpG methylomes from the murine frontal cortex, comprising 16 neuronal subtypes.
We subjected this data set to the standard scBS workflow: defining a set of genomic intervals, quantifying methylation in these intervals, and subjecting the resulting matrix to PCA.
We repeated this analysis several times, either using 100~kb tiles or VMRs, and either using methylation fractions or mean shrunken residuals as a measure of methylation. \todo{Is the following correct?} For both types of regions, the 100~kb tiles and the VMRs, we ranked the regions by the number of covering cells (i.e., cells with at least on read in the region) and kep only the top 1000 regions.

@@

...

@@

\todo{NOTE: Here, we could already do shrinkage and use $\frac{1}{|C_i|+1}\sum_i x_{ij}$ intead. But we don't do that, right?}

@@

Given a set, $\mathcal{V}=\{I^\text{v}_1,I^\text{v}_2,\dots\}$, of intervals corresponding to VMRs, we get a relative methylation fraction $r_{ij}$ for each VMR $I^\text{v}_i$ and each cell $j$ from Eq.\ ()\ref{avgres}). The matrix thus obtained can then be centered and used as input for a PCA. If we calculate the top $R$ principal components, we thus obtain for each cell $j$ an $R$-dimensional principal component vector $\mathbf{x}^\text{P}_j$. For any two cells $j$, $j'$, we use the Euclidean distance $\|\mathbf{x}^\text{P}_j - \mathbf{x}^\text{P}_{j'}\|$ as the measure of dissimilarity of the two cells. Thus, the matrix of PC scores can be used as input to dimension reduction methods like t-SNE or UMAP, and to clustering methods like the Louvain or Leiden algorithm, which requirte such a matrix as input to the approximate mnearest neighbor (ANN) finding algorithm that is their first step.

@@

\subsection{Nearest neighbor count score}
To assess performance of our methods, we processed mouse CpG-methylation data from \citet{luo2017single} with several method combinations and scored each result. The score evaluates the ability of each method combination to separate cell types in 15-dimensional PCA space, based on their CpG methylation. Each score varies between 0 and 1, where higher scores reflect better separation.

>>>>

\begin{figure}
    \begin{center}
    A\hspace{.9\columnwidth}~\\
    \includegraphics[width=.7\columnwidth]{part_leonie_git/leonie_plots/corr_methylation_fraction_4kbwindow.pdf} \\
    B\hspace{.9\columnwidth}~\\
    \includegraphics[width=.7\columnwidth]{part_leonie_git/leonie_plots/corr_shrunken_residual_4kbwindow.pdf} \\
    C\hspace{.9\columnwidth}~\\
    \includegraphics[width=.95\columnwidth]{part_leonie_git/leonie_plots/comparison_meth_RNA.pdf}
    \end{center}
    \caption{\small \textbf{Correlation of DNA methylation and gene expression}.
    \textbf{(A)} Correlation of each gene's RNA expression values to the the methylation fractions that were measured at promoters (y-axis), or to methylation fractions of the nearest VMR (x-axis). Promoters are defined as intervals \textpm2~kb around the TSS.
    \textbf{(B)} Same as in (A), but using mean shrunken residuals as a measure of methylation.
    \textbf{(C)} Mean smoothed methylation near the gene Ube2l6 (ENSMUSG00000027078, marked in yellow in A \& B). Cells are assigned to 4 groups based on Ube2l6 expression (group 0: cells with no Ube2l6 expression, groups 1-3: cells that express Ube2l6, divided into three equally large groups with group 3 having the highest expression). The mean methylation level of each group was smoothed with a tricube kernel of bandwidth 2000.
    Data from \citet{argelaguet2019gastru}.}
    \label{figure:correlation}
\end{figure}

@@

To assess the impact of our methods on dimensionality reduction and the ability to distinguish different (sub)types of cells, we used a data set from \citet{luo2017single}. This data set consists of 3377 single-cell CpG methylomes from the murine frontal cortex, comprising 16 neuronal subtypes.
We subjected this data set to the standard scBS workflow: defining a set of genomic intervals, quantifying methylation in these intervals, and subjecting the resulting matrix to PCA.
We repeated this analysis several times, either using 100~kb tiles or VMRs, and either using methylation fractions or mean shrunken residuals as a measure of methylation. \todo{Is the following correct?} For both types of regions, the 100~kb tiles and the VMRs, we ranked the regions by the number of covering cells (i.e., cells with at least on read in the region) and keep only the top 1000 regions.

@@

%Here we propose a number of new methods for the analysis of single cell methylation data and provide them to the research community as a command line tool called \texttt{scbs}.
%Instead of dividing the genome into coarse-grained tiles, we propose to discover variably-methylated regions (VMRs) in the data itself, which improves performance of common downstream applications.
%We also think that VMRs, due to their smaller size, are easier to interpret than  and furthermore provides a list of candidate regions.
%These regions may then be linked to individual genes, tested for celltype-specific methylation differences, or subjected to gene ontology or motif enrichment, depending on the research question at hand.
%We also propose a number of methods specifically tailored to the sparse nature of scBS data.
%First, we observe

%- propose a number of new methods for the analysis of scBS data
%- developed methods to analyse scBS data that go beyond simple genome tiling
% - instead we discover potential genomic regions of interest on the data itself, which improves performance of common downstream applications and aids interpretability.
% - VMRs can also be used as input for e.g. motif enrichment, go enrichment etc
%- devised a method to store genome-wide scBS data efficiently
%- propose a new measure of DNA methylation in a genomic interval (~ relative methylation)

%- implement them in Python and release them to the community as a Python package with a CLI
%-> no more custom scripts and manual data wrangling of 1000 text files required, instead every user can quickly go from raw data to count matrix
%- lowers the entry barrier for scBS data analysis and makes this new and exciting data type available to a greater audience

%whats missing?
%-

@@

%\todo{NOTE: Here, we could already do shrinkage and use $\frac{1}{|C_i|+1}\sum_i x_{ij}$ instead. But we don't do that, right?}

@@

Given a set, $\mathcal{V}=\{I^\text{v}_1,I^\text{v}_2,\dots\}$, of intervals corresponding to VMRs, we get a relative methylation fraction $r_{ij}$ for each VMR $I^\text{v}_i$ and each cell $j$ from Eq.\ ()\ref{avgres}). The matrix thus obtained can then be centered and used as input for a PCA. If we calculate the top $R$ principal components, we thus obtain for each cell $j$ an $R$-dimensional principal component vector $\mathbf{x}^\text{P}_j$. For any two cells $j$, $j'$, we use the Euclidean distance $\|\mathbf{x}^\text{P}_j - \mathbf{x}^\text{P}_{j'}\|$ as the measure of dissimilarity of the two cells.
Thus, the matrix of PC scores can be used as input to dimension reduction methods like t-SNE or UMAP, and to clustering methods like the Louvain or Leiden algorithm, which require such a matrix as input to the approximate mnearest neighbor (ANN) finding algorithm that is their first step.

@@

\subsection{Nearest neighbor count score}
To assess performance of our methods, we processed mouse CpG methylation data from \citet{luo2017single} with several method combinations and scored each result. The score evaluates the ability of each method combination to separate cell types in 15-dimensional PCA space, based on their CpG methylation. Each score varies between 0 and 1, where higher scores reflect better separation.
